---
title: "Fish-Ecol Org"
author: "Faúndez-Baez, P., Hartline, N., Mayorga-Henao, J., Villaseñor-Derbez, J.C."
output:
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
---

##Types of data

We have the following data:

- IUCN: Data on endangerment of fish species (Chondricthyes). The data will include trend, global status, country, habitat (freshwater or marine), and other species details. Some of these data sets will need to be joined.
- Hunt-data
    - This data set contains estimates of population abundance for 5 species of reef fish and their associated recruitment habitat. The data was collected in 2002 and 2014 at two lagoons in French Polynesia that have contrasting anthropogenic pressures. 
 
## Data and metadata standards
 
Data comes from the International Union for Conservation of Nature: IUCN Red List of Threatened Species
http://www.iucnredlist.org
The IUCN Red List of Threatened Species is widely recognized as the most comprehensive, objective global approach for evaluating the conservation status of plant and animal species. From its small beginning, the IUCN Red List has grown in size and complexity and now plays an increasingly prominent role in guiding conservation activities of governments, NGOs and scientific institutions. The introduction in 1994 of a scientifically rigorous approach to determine risks of extinction that is applicable to all species, has become a world standard. In order to produce the IUCN Red List of Threatened Species, the IUCN Species Programme working with the IUCN Survival Commission (SSC) and with members of IUCN draws on and mobilizes a network of scientists and partner organizations working in almost every country in the world, who collectively hold what is likely the most complete scientific knowledge base on the biology and conservation status of species.
 
## Policies for access and sharing

All data, program code and results are freely available via the service GitHub https://github.com/fish-ecol

Data should be cited as: 

Faúndez-Baez, P., Hartline, N., Mayorga-Henao, J., Villaseñor-Derbez, J.C. "titlle"

and

IUCN 2015. The IUCN Red List of Threatened Species. Version 2015-4. <http://www.iucnredlist.org>. Downloaded on 05 February 2016.


## Plans for archiving and preservation of access

Are we thinking into submit our data to a national data center, publication in a widely available scientific journal, book or website? If not, we should take off this point.


## Data question(s)


```{r, echo=FALSE, warning=FALSE}

suppressWarnings(suppressMessages(library(readr)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(tidyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(knitr)))
suppressWarnings(suppressMessages(library(rworldmap)))
suppressWarnings(suppressMessages(library(ggmap)))

```


### IUCN Data for Elasmobranchs

```{r,echo=FALSE}
elasmo <- read_csv("Data/Elasmo.csv")

SN=length(unique(elasmo$`Scientific Name`))
C=length(unique(elasmo$Class))
O=length(unique(elasmo$Order))
Fa=length(unique(elasmo$Family))
```

Some descriptive stuff for us to know what we have.

**Table 1-** Taxonomic summary of the IUCN data.

Classes | Orders | Families  | Species
--------|--------|-----------|--------
`r C`   | `r O`  |  `r Fa`   | `r SN`


```{r, echo=FALSE, fig.cap="Categories for ech Order"}

ggplot(data=elasmo, aes(Order))+
  geom_bar(aes(fill=Category))+
  scale_fill_brewer(palette="Set1")+
  theme_bw()+
  coord_flip()

```

```{r, echo=FALSE, fig.cap="Number of cases by Category"}

ggplot(data=elasmo, aes(Category))+
  geom_bar(aes(fill=Category))+
  scale_fill_brewer(palette="Set1")+
  theme_bw()
```

```{r, echo=FALSE, fig.cap="Changes in category though years"}

elasmo2=elasmo %>%
  group_by(`Modified Year`, Category) %>%
  summarize(Count=n())

ggplot(data=elasmo2, aes(`Modified Year`, Count))+
  geom_line(aes(col=Category))+
  scale_color_brewer(palette="Set1")+
  theme_bw()
```

### Option 4 Hunter-data

Exploratory analysis of this dataset. 
```{r,results = 'asis', echo=FALSE}
fish <- read_csv("Data/hunter-data.csv")
kable(fish)
```

#### What is the change in abundances of adults at each site?

```{r}
change<- fish %>% 
  select(Species, Adults, Island, year) %>%  
  group_by(Species, Island) %>% 
  spread(year, Adults) %>% 
  mutate(Change =100*(`2014`-`2002`)/`2002`) 
kable(change)
```

```{r, echo=FALSE, fig.cap="Relative change in mean density of adults (individuals per 100m2 of reef)"}
ggplot(change, aes(Species, Change)) +
  geom_bar(stat = "identity", aes(fill = Island),position=position_dodge(), width=.5)+
  scale_fill_brewer(palette="Set1")+
  theme_bw()+
  geom_hline(yintercept = 0)+
  theme(panel.grid.minor.x=element_blank(), panel.grid.major.x=element_blank())+
  theme(panel.grid.minor.y=element_blank(), panel.grid.major.y=element_blank())+
  labs(x='Species',y='% Change')
```

```{r,echo=FALSE,fig.width=4, fig.height=6}
setwd("./Data")
Taxonomic=read.csv("Fish Red List Taxonomic and Status.csv")
#Note that some of the scientific names are a little bit funky in these files. They seem to have not too much information on them and as such have weird characters that may confuse any code used for joining and such.

#convert low count categories to closest related cateogry (consolidating categories-- we can also just remove them). #note that grep below finds the index of LR/cd OR LR/lc.
Taxonomic$Category[grep("LR/cd|LR/lc",Taxonomic$Category)]="LC" 
Taxonomic$Category[grep("LR/nt",Taxonomic$Category)]="NT"
Taxonomic$Category[grep("EW",Taxonomic$Category)]="EX"
Taxonomic$Category=factor(Taxonomic$Category,levels=c("EX","CR","EN","VU","NT","LC","DD")) 

library(RColorBrewer)
library(ggplot2)

SummaryTable=data.frame(table(Taxonomic$Category,Taxonomic$Class))
names(SummaryTable)=c("GlobalStatus","Class","Count")

colours=brewer.pal(name="OrRd", n=nlevels(SummaryTable$GlobalStatus))
names(colours)=rev(levels(SummaryTable$GlobalStatus))

CategoryGraph=list()
for(i in 1:nlevels(SummaryTable$Class)){
  temp=SummaryTable[SummaryTable$Class==levels(SummaryTable$Class)[i],]
  CategoryGraph[[i]]=ggplot(data=temp,aes(x=Class,y=Count,fill=GlobalStatus,order=GlobalStatus,width=.3))+
    geom_bar(,stat="identity")+
    guides(fill=guide_legend(reverse=TRUE))+
    scale_fill_manual(values=colours)+
    ylab("Number of Species")+
    ggtitle("IUCN Threat Categories")+
    theme_bw()+
    scale_x_discrete(labels=c("ACTINOPTERYGII"="Actinopterygii","CEPHALASPIDOMORPHI"="Cephalaspidomorphi","CHONDRICHTHYES"="Chondrichthyes","MYXINI"="Myxini","SARCOPTERYGII"="Sarcopterygii"))+
    geom_bar(stat="identity",color="black",show.legend=FALSE)+
    xlab(NULL)+
    theme(legend.background = element_rect(colour = "black"),plot.title = element_text(size = 17, vjust = 2),axis.text=element_text(size=17),axis.title=element_text(size=15,vjust = 1),legend.title=element_blank(),legend.text=element_text(size=13))
}
CategoryGraph

#I spent a little bit of time getting these put together. Having them on the same graph doesn't work since Actinopterygii is so much larger than the other fish classes. I was looking into getting them laid out horizontally (something like facet_grid) with unique axes for each but only a single title and one legend. It looks like working with Actinopterygii and Chondrichthyes is our best option if we want to do more than Chondrichthyes. The next step that might be good is to use 'taxize' to obtain encyclopedia of life (EOL) information on the weight of the species in chondrichthyes.

```



